---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# eggla

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/eggla)](https://CRAN.R-project.org/package=eggla)
[![R build status](https://github.com/mcanouil/eggla/workflows/R-CMD-check/badge.svg)](https://github.com/mcanouil/eggla/actions)
<!-- badges: end -->

Tools for longitudinal analysis within the EGG (Early Growth Genetics) Consortium.

## Installation

```{r, eval = FALSE}
# Install insane from CRAN:
install.packages("eggla")

# Or the the development version from GitHub:
# install.packages("remotes")
remotes::install_github("mcanouil/eggla")
```

## Hands on **eggla**

```{r}
library(eggla)
library(data.table)
library(ggplot2)
library(patchwork)
library(broom.mixed)
```

### Data

**eggla** includes a small dataset with 1050 rows and 6 variables.

* `ID` (`character`): ID using three digits.
* `age` (`numeric`): age in years.
* `sex` (`integer`): sex with 1: male and 0: female.
* `weight` (`numeric`): weight in kilograms.
* `height` (`integer`): height in centimetres.
* `bmi` (`numeric`): Body Mass Index in kilograms per quare metre.

```{r}
data("bmigrowth")
bmigrowth[bmigrowth[["ID"]] == "001", ]
```


### Modelling

**eggla** includes three models, namely `cubic_slope`, `linear_splines` and `cubic_splines`, all implemented with the R package `nlme`.

* `x` and `y`, the variables as character with possible transformation.
* `data`, the data in which the variables are.
* `method`, the method or model to be performed.
* `knots`, the knots for "splines" methods.
    Default is the following:
    ```{r, eval = FALSE}
    knots = list(
      "cubic_slope" = NULL,
      "linear_splines" = c(5.5, 11),
      "cubic_splines" = c(2, 8, 12)
    )[[method]]
    ```

```{r}
ls_mod <- time_model(
  x = "age", 
  y = "log(bmi)", 
  data = bmigrowth[bmigrowth$sex == 0, ], 
  method = "linear_splines"
)
```

The results of `time_model` is an `lme` object, on which any usual methods can be applied, *e.g., `summary()` or `coefficients()`.

```{r}
class(ls_mod)
tidy(ls_mod)
```

### Residuals

Different plot are available for model diagnostic, using the residuals mostly.

```{r}
plot_residuals(
  x = "age", 
  y = "log(bmi)", 
  fit = ls_mod, 
  variables_unit = list(age = "years", bmi = "kg/m²")
) + 
  plot_annotation(
    title = "LINEAR SPLINES - BMI - Female", 
    tag_levels = "A"
  )
```

### Predicted Average Slopes

```{r}
ls_pred_slopes <- predict_average_slopes(
  fit = ls_mod,
  method = "linear_splines",
  period = c(0, 0.5, 1.5, 5, 6, 10, 12, 17)#,
  # knots = list(
  #   "cubic_slope" = NULL,
  #   "linear_splines" = c(5.5, 11),
  #   "cubic_splines" = c(2, 8, 12)
  # )[[method]]
)
head(ls_pred_slopes)
```

```{r}
ggplot(
  data = melt(
    data = setDT(ls_pred_slopes), 
    id.vars = c("ID"), 
    measure.vars = patterns("^slope_"),
    variable.name = "period_interval", 
    value.name = "slope"
  )[, 
    period_interval := factor(
      x = gsub("slope_", "", period_interval), 
      levels = gsub("slope_", "", unique(period_interval))
    )
  ]
) +
  aes(x = slope) +
  geom_histogram(bins = 30) +
  scale_y_continuous(expand = expansion(c(0, 0.1))) +
  facet_grid(cols = vars(period_interval), scales = "free") +
  labs(x = "Predicted Slope", y = "Count")
```

### Area Under The Curves

```{r}
ls_auc <- compute_auc(
  fit = ls_mod,
  method = "linear_splines",
  period = c(0, 0.5, 1.5, 5, 6, 10, 12, 17)#,
  # knots = list(
  #   "cubic_slope" = NULL,
  #   "linear_splines" = c(5.5, 11),
  #   "cubic_splines" = c(2, 8, 12)
  # )[[method]]
)
head(ls_auc)
```

```{r}
ggplot(
  data = melt(
    data = setDT(ls_auc), 
    id.vars = "ID", 
    measure.vars = patterns("^auc_"),
    variable.name = "period_interval", 
    value.name = "auc"
  )[, 
    period_interval := factor(
      x = gsub("auc_", "", period_interval), 
      levels = gsub("auc_", "", unique(period_interval))
    )
  ]
) +
  aes(x = auc) +
  geom_histogram(bins = 30) +
  scale_y_continuous(expand = expansion(c(0, 0.1))) +
  facet_grid(cols = vars(period_interval), scales = "free") +
  labs(x = "Area Under The Curve (AUC)", y = "Count")
```

### Render Analyses As Rmarkdown

```{r}
library(eggla)
data("bmigrowth")

file.copy(
  from = system.file("rmarkdown", "templates", "eggla", "skeleton", "skeleton.Rmd", package = "eggla"),
  to = file.path(tempdir(), "eggla.Rmd"),
  overwrite = TRUE
)

rmarkdown::render(
  input = file.path(tempdir(), "eggla.Rmd"),
  params = list(
    cohort_name =  "BMI Growth (Demo)",
    phenotypes = bmigrowth,
    trait = "bmi",
    output_directory = tempdir()
  )
)

file.copy(file.path(tempdir(), "eggla.html"), "eggla.html")
```

<!-- ```{r, eval = FALSE} -->
<!-- run_eggla( -->
<!--   phenotypes = bmigrowth, -->
<!--   trait = "bmi", -->
<!--   cohort_name = "cohort", -->
<!--   output_directory = tempdir() -->
<!-- ) -->
<!-- ``` -->

