---
title: "Get Started with 'eggla'"
description: |
  How-to run the 'eggla' workflow from quality-control to genome-wide association study.
author: "Mickaël Canouil, *Ph.D.* (<mickael.canouil@cnrs.fr>)"
output: 
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
    toc_depth: 3
vignette: |
  %\VignetteIndexEntry{Get Started with 'eggla'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Run Using Docker

## Create `run_eggla.R` script

Copy and edit the following code to a new file (_e.g._, `/home/mcanouil/eggla/run_eggla.R`) on the server that will run the analysis with the appropriate parameters.

```{r, eval = FALSE}
library(eggla)
library(data.table)
res <- try(run_eggla(
  data = fread("/home/mcanouil/eggla/bmigrowth.csv"), # to be changed with the path of the file containing the data
  id_variable = "ID",
  age_days_variable = NULL, # computed based on "age_years_variable" if not provided. Only used for QC.
  age_years_variable = "age",
  weight_kilograms_variable = "weight",
  height_centimetres_variable = "height",
  sex_variable = "sex",
  covariates = NULL,
  male_coded_zero = FALSE,
  random_complexity = "auto",
  use_ar1 = FALSE,
  parallel = FALSE, # to parallelise Daymont QC
  parallel_n_chunks = 1, # to parallelise Daymont QC
  working_directory = "."
))
if (inherits(res, "try-error")) unlink(wd, recursive = TRUE)
```

## Run the script using a Docker container

Donwload the (latest) Docker image or a specific version from the [GitHub Registry](https://github.com/users/mcanouil/packages/container/package/eggla) and run it with the following command:

```{bash, eval = FALSE}
docker run \
  -it --rm \
  --volume /home/mcanouil/eggla:/tmp/eggla \
  ghcr.io/mcanouil/eggla:latest Rscript /tmp/eggla/run_eggla.R
```

_Note_: the left-hand side (LHS) of `/home/mcanouil/eggla:/tmp/eggla` is the directory on the server, the right-hand side is how LHS directory will appear within the Docker container.

## Retrieve the outputs

```
/home/mcanouil/eggla/
├── 2021-11-23-female.zip
└── 2021-11-23-male.zip
```

## Create `run_eggla_gwas.R` Script

Copy and edit the following code to a new file (_e.g._, `/home/mcanouil/eggla/run_eggla_gwas.R`) on the server that will run the analysis with the appropriate parameters.

```{r, eval = FALSE}
working_directory <- "/home/mcanouil/eggla"
setwd(working_directory)
library(eggla)
chr_in_parallel <- FALSE # or 11 to run 11 chromosomes in 11 processes/cores
if (!isFALSE(chr_in_parallel)) {
  renv::install(c("future", "future.apply"))
  future::plan("multicore", workers = min(chr_in_parallel, future::availableCores()))
}
do_eggla_gwas(
  data = "./bmigrowth.csv",
  results_zip = list.files(path = ".", pattern = "\\.zip", full.names = TRUE),
  id_column = "ID",
  traits = c("slope_.*", "auc_.*"),
  covariates = c("sex"),
  vcfs = list.files(path = file.path(".", "vcf"), pattern = "\\.vcf$|\\.vcf.gz$", full.names = TRUE),
  vep = NULL,
  path = ".",
  bin_path = list(
    bcftools = "/usr/bin/bcftools",
    plink2 = "https://s3.amazonaws.com/plink2-assets/plink2_linux_x86_64_20220410.zip" # check URL before using
  ),
  threads = 1
)
```

## Run the script using a Docker container

Donwload the (latest) Docker image or a specific version from the [GitHub Registry](https://github.com/users/mcanouil/packages/container/package/eggla) and run it with the following command:

```{bash, eval = FALSE}
docker run \
  -it --rm \
  --volume /home/mcanouil/eggla:/tmp/eggla \
  ghcr.io/mcanouil/eggla:latest Rscript /tmp/eggla/run_eggla_gwas.R
```

_Note_: the left-hand side (LHS) of `/home/mcanouil/eggla:/tmp/eggla` is the directory on the server, the right-hand side is how LHS directory will appear within the Docker container.

## Retrieve the outputs

```
/home/mcanouil/eggla/
├── gwas_software.txt
└── gwas.csv.gz
```

# Run Non-Interactively (In Bash/Shell)

## Create `run_eggla.sh` script

Copy and edit the following code to a new file (_e.g._, `run_eggla.sh`) on the server that will run the analysis with the appropriate parameters.

### `renv` (recommended)

```{bash, eval = FALSE}
#!/bin/bash

home_analysis="/home/mcanouil/eggla" # to be changed to the folder in which "egg_analysis" is to be performed

mkdir $home_analysis 

cd $home_analysis || exit

Rscript <<EOF
temp_library <- file.path(".", "R")
dir.create(temp_library, recursive = TRUE)
.libPaths(temp_library)
install.packages("renv", lib = temp_library, repos = "http://cloud.r-project.org")
library("renv")
renv::init(bare = TRUE, settings = list(use.cache = FALSE))
utils::download.file(
    url = "https://raw.githubusercontent.com/mcanouil/eggla/v0.9.1/inst/setup/renv.lock",
    destfile = "renv.lock"
  )
renv::restore(lockfile = "renv.lock")
renv::install("mcanouil/eggla@v0.9.1")
unlink(temp_library, recursive = TRUE)
EOF

Rscript <<EOF
library(eggla)
library(data.table)
res <- try(run_eggla(
  data = fread("bmigrowth.csv"), # to be changed with the path of the file containing the data
  id_variable = "ID",
  age_days_variable = NULL, # computed based on "age_years_variable" if not provided. Only used for QC.
  age_years_variable = "age",
  weight_kilograms_variable = "weight",
  height_centimetres_variable = "height",
  sex_variable = "sex",
  covariates = NULL,
  male_coded_zero = FALSE,
  random_complexity = "auto",
  use_ar1 = FALSE,
  parallel = FALSE, # to parallelise Daymont QC
  parallel_n_chunks = 1, # to parallelise Daymont QC
  working_directory = "."
))
if (inherits(res, "try-error")) unlink(wd, recursive = TRUE)
EOF
```

### `pak` (faster)

```{bash, eval = FALSE}
#!/bin/bash

home_analysis="/home/mcanouil/eggla" # to be changed to the folder in which "egg_analysis" is to be performed

mkdir $home_analysis 

cd $home_analysis || exit

Rscript <<EOF
temp_library <- file.path(".", "R")
dir.create(temp_library, recursive = TRUE)
.libPaths(temp_library)
utils::install.packages(
  pkgs = "pak",
  lib = temp_library,
  repos = sprintf(
    "https://r-lib.github.io/p/pak/stable/%s/%s/%s",
    .Platform[["pkgType"]], R.Version()[["os"]], R.Version()[["arch"]]
  )
)
library(pak)
utils::download.file(
  url = "https://raw.githubusercontent.com/mcanouil/eggla/v0.9.1/inst/setup/pkg.lock",
  destfile = "pkg.lock"
)
lockfile_install(lockfile = "pkg.lock", lib = temp_library)
pkg_install("mcanouil/eggla@v0.9.1", lib = temp_library, upgrade = FALSE, dependencies = FALSE)
EOF

Rscript <<EOF
library(eggla)
library(data.table)
res <- try(run_eggla(
  data = fread("bmigrowth.csv"), # to be changed with the path of the file containing the data
  id_variable = "ID",
  age_days_variable = NULL, # computed based on "age_years_variable" if not provided. Only used for QC.
  age_years_variable = "age",
  weight_kilograms_variable = "weight",
  height_centimetres_variable = "height",
  sex_variable = "sex",
  covariates = NULL,
  male_coded_zero = FALSE,
  random_complexity = "auto",
  use_ar1 = FALSE,
  parallel = FALSE, # to parallelise Daymont QC
  parallel_n_chunks = 1, # to parallelise Daymont QC
  working_directory = "."
))
if (inherits(res, "try-error")) unlink(wd, recursive = TRUE)
EOF
```

## Run the analysis

```{bash, eval = FALSE}
bash run_eggla.sh
```

## Retrieve the outputs

```
/home/mcanouil/eggla/
├── 2021-11-23-female.zip
└── 2021-11-23-male.zip
```

## Run the GWASs

Run GWASs on slopes and AUCs for male and female.

```{bash, eval = FALSE}
#!/bin/bash

home_analysis="/home/mcanouil/eggla" # to be changed to the folder in which "egg_analysis" is to be performed

mkdir $home_analysis

cd $home_analysis || exit

Rscript <<EOF
library(eggla)
chr_in_parallel <- FALSE # or 11 to run 11 chromosomes in 11 processes/cores
if (!isFALSE(chr_in_parallel)) {
  renv::install(c("future", "future.apply"))
  future::plan("multicore", workers = min(chr_in_parallel, future::availableCores()))
}
do_eggla_gwas(
  data = "bmigrowth.csv",
  results_zip = list.files(path = ".", pattern = "\\\\.zip", full.names = TRUE),
  id_column = "ID",
  traits = c("slope_.*", "auc_.*"),
  covariates = c("sex"),
  vcfs = list.files(path = file.path(".", "vcf"), pattern = "\\\\.vcf$|\\\\.vcf.gz$", full.names = TRUE),
  vep = NULL,
  path = ".",
  bin_path = list(
    bcftools = "/usr/bin/bcftools",
    plink2 = "https://s3.amazonaws.com/plink2-assets/plink2_linux_x86_64_20220410.zip" # check URL before using
  ),
  threads = 1
)
EOF
```

## Retrieve the outputs

```
/home/mcanouil/eggla/
├── gwas_software.txt
└── gwas.csv.gz
```

# Run Interactively In Bash/R

## Create the working directory

```{bash, eval = FALSE}
home_analysis="/home/mcanouil/eggla" # to be changed to the folder in which "egg_analysis" is to be performed

mkdir $home_analysis 

cd $home_analysis || exit
```

## Start R and setup environement

Start R and setup the working directory using `renv` (recommended) or `pak` (faster) to restore predefined version of packages.

### `renv` (recommended)

```{r, eval = FALSE}
temp_library <- file.path(".", "R")
dir.create(temp_library, recursive = TRUE)
.libPaths(temp_library)
install.packages("renv", lib = temp_library, repos = "http://cloud.r-project.org")
library("renv")
renv::init(bare = TRUE, settings = list(use.cache = FALSE))
utils::download.file(
  url = "https://raw.githubusercontent.com/mcanouil/eggla/v0.9.1/inst/setup/renv.lock",
  destfile = "renv.lock"
)
renv::restore(lockfile = "renv.lock")
renv::install("mcanouil/eggla@v0.9.1")
unlink(temp_library, recursive = TRUE)
```

### `pak` (faster)

```{r, eval = FALSE}
temp_library <- file.path(".", "R")
dir.create(temp_library, recursive = TRUE)
.libPaths(temp_library)
utils::install.packages(
  pkgs = "pak",
  lib = temp_library,
  repos = sprintf(
    "https://r-lib.github.io/p/pak/stable/%s/%s/%s",
    .Platform[["pkgType"]], R.Version()[["os"]], R.Version()[["arch"]]
  )
)
library(pak)
utils::download.file(
  url = "https://raw.githubusercontent.com/mcanouil/eggla/v0.9.1/inst/setup/pkg.lock",
  destfile = "pkg.lock"
)
lockfile_install(lockfile = "pkg.lock", lib = temp_library)
pkg_install("mcanouil/eggla@0.9.1", lib = temp_library, upgrade = FALSE, dependencies = FALSE)
```


## Run the analysis

Restart R before running the analysis.

```{r, eval = FALSE}
# setwd("/home/mcanouil/eggla") # already set if step 1 was done
library(eggla)
library(data.table)
res <- try(
  run_eggla(
    data = fread("bmigrowth.csv"),
    id_variable = "ID",
    age_days_variable = NULL,
    age_years_variable = "age",
    weight_kilograms_variable = "weight",
    height_centimetres_variable = "height",
    sex_variable = "sex",
    covariates = NULL,
    male_coded_zero = FALSE,
    random_complexity = "auto",
    use_ar1 = FALSE,
    parallel = FALSE,
    parallel_n_chunks = 1,
    working_directory = "."
  )
)
if (inherits(res, "try-error")) unlink(wd, recursive = TRUE)
```

## Retrieve the outputs

```
/home/mcanouil/eggla/
├── 2021-11-23-female.zip
└── 2021-11-23-male.zip
```

## Run the GWASs

Run the GWASs on slopes and AUCs for male and female.

```{r, eval = FALSE}
# setwd("/home/mcanouil/eggla") # already set if step 1 was done
library(eggla)
chr_in_parallel <- FALSE # or 11 to run 11 chromosomes in 11 processes/cores
if (!isFALSE(chr_in_parallel)) {
  renv::install(c("future", "future.apply"))
  future::plan("multicore", workers = min(chr_in_parallel, future::availableCores()))
}
do_eggla_gwas(
  data = "bmigrowth.csv",
  results_zip = list.files(path = ".", pattern = "\\.zip", full.names = TRUE),
  id_column = "ID",
  traits = c("slope_.*", "auc_.*"),
  covariates = c("sex"),
  vcfs = list.files(
    path = file.path(".", "vcf"),
    pattern = "\\.vcf$|\\.vcf.gz$",
    full.names = TRUE
  ),
  vep = NULL,
  path = ".",
  bin_path = list(
    bcftools = "/usr/bin/bcftools",
    plink2 = "https://s3.amazonaws.com/plink2-assets/plink2_linux_x86_64_20220410.zip" # check URL before using
  ),
  threads = 1
)
```

## Retrieve the outputs

```
/home/mcanouil/eggla/
├── gwas_software.txt
└── gwas.csv.gz
```
