---
title: "Run 'eggla' Workflow"
author: "Mickaël Canouil, *Ph.D.* (<***REMOVED***>)"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Run Non-Interactively

1. Copy and edit the following code to a new file (_e.g._, `run_eggla.sh`) on the server that will run the analysis with the appropriate parameters.

  - `renv` (recommended):

    ```{bash, eval = FALSE}
    #!/bin/bash

    home_analysis="/tmp/egg_analysis" # to be changed to the folder in which "egg_analysis" is to be performed

    mkdir $home_analysis 

    cd $home_analysis

    Rscript \
      -e 'wd <- "/tmp/egg_analysis"' \
      -e 'temp_library <- file.path(wd, "R")' \
      -e 'dir.create(temp_library, recursive = TRUE)' \
      -e '.libPaths(temp_library)' \
      -e 'install.packages("renv", lib = temp_library, repos = "http://cloud.r-project.org")' \
      -e 'library("renv")' \
      -e 'renv::init(bare = TRUE, settings = list(use.cache = FALSE))' \
      -e 'renv::restore(lockfile = "https://raw.githubusercontent.com/mcanouil/eggla/main/inst/setup/renv.lock")' \
      -e 'renv::install("mcanouil/eggla@v0.5.0")' \
      -e 'unlink(temp_library, recursive = TRUE)'

    Rscript \
      -e 'wd <- "/tmp/egg_analysis"' \
      -e 'setwd(wd)' \
      -e 'library(eggla)' \
      -e 'library(data.table)' \
      -e 'res <- try(run_eggla(
        data = fread("/tmp/bmigrowth.csv"), # to be changed with the path of the file containing the data
        id_variable = "ID",
        age_days_variable = NULL, # computed based on "age_years_variable" if not provided. Only used for QC.
        age_years_variable = "age", 
        weight_kilograms_variable = "weight",
        height_centimetres_variable = "height",
        sex_variable = "sex",
        covariates = NULL,
        male_coded_zero = FALSE,
        parallel = FALSE, # to parallelise Daymont QC
        parallel_n_chunks = 1, # to parallelise Daymont QC
        working_directory = wd # or in that case "/tmp/egg_analysis"
      ))' \
      -e 'if (inherits(res, "try-error")) unlink(wd, recursive = TRUE)'
    ```

  - `pak` (faster):

    ```{bash, eval = FALSE}
    #!/bin/bash

    home_analysis="/tmp/egg_analysis" # to be changed to the folder in which "egg_analysis" is to be performed

    mkdir $home_analysis 

    cd $home_analysis

    Rscript \
      -e 'wd <- "/tmp/egg_analysis"' \
      -e 'temp_library <- file.path(wd, "R")' \
      -e 'dir.create(temp_library, recursive = TRUE)' \
      -e '.libPaths(temp_library)' \
      -e 'install.packages("pak", lib = temp_library, repos = "https://r-lib.github.io/p/pak/devel/")' \
      -e 'library(pak)' \
      -e 'lockfile_install(
        lockfile = "https://raw.githubusercontent.com/mcanouil/eggla/main/inst/setup/pkg.lock",
        lib = temp_library
      )' \
      -e 'pkg_install("mcanouil/eggla@v0.5.0", lib = temp_library, upgrade = FALSE, dependencies = FALSE)'
    
    Rscript \
      -e 'wd <- "/tmp/egg_analysis"' \
      -e 'setwd(wd)' \
      -e 'library(eggla)' \
      -e 'library(data.table)' \
      -e 'res <- try(run_eggla(
        data = fread("/tmp/bmigrowth.csv"), # to be changed with the path of the file containing the data
        id_variable = "ID",
        age_days_variable = NULL, # computed based on "age_years_variable" if not provided. Only used for QC.
        age_years_variable = "age", 
        weight_kilograms_variable = "weight",
        height_centimetres_variable = "height",
        sex_variable = "sex",
        covariates = NULL,
        male_coded_zero = FALSE,
        parallel = FALSE, # to parallelise Daymont QC
        parallel_n_chunks = 1, # to parallelise Daymont QC
        working_directory = wd # or in that case "/tmp/egg_analysis"
      ))' \
      -e 'if (inherits(res, "try-error")) unlink(wd, recursive = TRUE)'
    ```

2.  Run the analysis in bash

    ```{bash, eval = FALSE}
    bash run_eggla.sh
    ```

3. Retrieve the two archives

    ```
    /tmp/egg_analysis/
    ├── 2021-11-23-female.zip
    └── 2021-11-23-male.zip
    ```

# Run Interactively

1. Create the working directory

    ```{bash, eval = FALSE}
    home_analysis="/tmp/egg_analysis" # to be changed to the folder in which "egg_analysis" is to be performed
    
    mkdir $home_analysis 

    cd $home_analysis
    ```

2. Start R and setup working directory using `renv` (recommended) or `pak` (faster) to restore predefined version of packages

  - `renv` (recommended):

    ```{r, eval = FALSE}
    wd <- "/tmp/egg_analysis"
    temp_library <- file.path(wd, "R")
    dir.create(temp_library, recursive = TRUE)
    .libPaths(temp_library)
    install.packages("renv", lib = temp_library, repos = "http://cloud.r-project.org")
    library("renv")
    renv::init(bare = TRUE, settings = list(use.cache = FALSE))
    renv::restore(lockfile = "https://raw.githubusercontent.com/mcanouil/eggla/main/inst/setup/renv.lock")
    renv::install("mcanouil/eggla@v0.5.0")
    unlink(temp_library, recursive = TRUE)
    ```

  - `pak` (faster):

    ```{r, eval = FALSE}
    wd <- "/tmp/egg_analysis"
    temp_library <- file.path(wd, "R")
    dir.create(temp_library, recursive = TRUE)
    .libPaths(temp_library)
    install.packages("pak", lib = temp_library, repos = "https://r-lib.github.io/p/pak/devel/")
    library(pak)
    lockfile_install(
      lockfile = "https://raw.githubusercontent.com/mcanouil/eggla/main/inst/setup/pkg.lock",
      lib = temp_library
    )
    pkg_install("mcanouil/eggla@0.5.0", lib = temp_library, upgrade = FALSE, dependencies = FALSE)
    ```

3. Restart R

4. Run the analysis

    ```{r, eval = FALSE}
    setwd(wd)
    library(eggla)
    library(data.table)
    res <- try(
      run_eggla(
        data = fread("/tmp/bmigrowth.csv"),
        id_variable = "ID",
        age_days_variable = NULL,
        age_years_variable = "age",
        weight_kilograms_variable = "weight",
        height_centimetres_variable = "height",
        sex_variable = "sex",
        covariates = NULL,
        male_coded_zero = FALSE,
        parallel = FALSE,
        parallel_n_chunks = 1,
        working_directory = wd
      )
    )
    if (inherits(res, "try-error")) unlink(wd, recursive = TRUE)
    ```

5. Retrieve the two archives

    ```
    /tmp/egg_analysis/
    ├── 2021-11-23-female.zip
    └── 2021-11-23-male.zip
    ```
