---
title: "Run The Cubic Splines (Random Cubic/Linear Splines) Analysis"
author: "Mickaël Canouil, *Ph.D.* (<***REMOVED***>)"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup

```{r}
library(eggla)
library(growthcleanr)
library(broom.mixed)
library(data.table, quietly = TRUE)

# Setup for plots
library(ggplot2, quietly = TRUE)
library(patchwork)
theme_set(theme_minimal())
options(
  ggplot2.discrete.colour = function(...) scale_colour_viridis_d(..., begin = 0.15, end = 0.85),
  ggplot2.discrete.fill = function(...) scale_fill_viridis_d(..., begin = 0.15, end = 0.85),
  ggplot2.continuous.colour = function(...) scale_colour_viridis_c(..., begin = 0.15, end = 0.85),
  ggplot2.continuous.fill = function(...) scale_fill_viridis_c(..., begin = 0.15, end = 0.85)
)
```

# Data

**eggla** includes a small dataset with 1050 rows and 6 variables.

* `ID` (`character`): ID using three digits.
* `age` (`numeric`): age in years.
* `sex` (`integer`): sex with 1: male and 0: female.
* `weight` (`numeric`): weight in kilograms.
* `height` (`integer`): height in centimetres.
* `bmi` (`numeric`): Body Mass Index in kilograms per quare metre.

  ```{r}
  data("bmigrowth")
  bmigrowth <- as.data.table(bmigrowth)
  ```
  ```{r}
  bmigrowth[bmigrowth[["ID"]] == "001"]
  ```
  ```{r}
  ggplot(data = bmigrowth) +
    aes(x = age, y = bmi, colour = factor(ID)) +
    geom_path(na.rm = TRUE, alpha = 0.25) +
    geom_point(size = 0.5, na.rm = TRUE, alpha = 0.25) +
    stat_smooth(
      method = "gam",
      formula = y ~ s(x, bs = "cr"),
      linetype = 1,
      colour = "firebrick",
      se = FALSE
    ) +
    theme(legend.position = "none") +
    labs(x = "AGE (years)", y = "BMI (kg/m\u00B2)") +
    facet_grid(
      cols = vars(sex),
      margins = TRUE,
      labeller = labeller(
        .cols = function(x) {
          c("0" = "FEMALE", "1" = "MALE", "2" = "FEMALE", "(all)" = "ALL")[x]
        }
      )
    )
  ```

## With Daymont's QC

```{r}
pheno_dt <- bmigrowth[
  j = `:=`(
    "agedays" = floor(age * 365.25), # convert to age in days and as integers ...
    "WEIGHTKG" = as.numeric(weight),
    "HEIGHTCM" = as.numeric(height)
  )
][
  j = `:=`(# recode sex with Male = 0 and Female = 1...
    "sex_daymont" = c("0" = "1", "1" = "0")[as.character(sex)]
  )
]

pheno_dt_long <- melt(
  data = pheno_dt,
  id.vars = c("ID", "age", "sex", "agedays", "sex_daymont"),
  measure.vars = c("WEIGHTKG", "HEIGHTCM"),
  variable.name = "param",
  value.name = "measurement"
)[
  j = clean := cleangrowth(# Daymont's QC from 'growthcleanr'
    subjid = ID,
    param = param,
    agedays = agedays,
    sex = sex_daymont,
    measurement = measurement,
    quietly = TRUE
  )
]
pheno_dt_clean <- dcast(
  data = pheno_dt_long[clean %in% "Include"], # Exclude all flags
  formula = ... ~ param,
  value.var = "measurement"
)[
  j = "bmi" := WEIGHTKG / (HEIGHTCM / 100)^2 # recompute bmi based using QCed variables
][
  !is.na(bmi) # exclude missing BMI related to measurements exclusion
]
```

## Without Daymont's QC

```{r, eval = FALSE}
pheno_dt_clean <- bmigrowth
```

# Modelling Female

```{r}
pheno_dt_female <- pheno_dt_clean[sex_daymont == 1]
```

```{r}
res <- egg_model(
  formula = log(bmi) ~ age, # + covariates, e.g., `log(bmi) ~ age + covariates`
  data = pheno_dt_female,
  id_var = "ID",
  random_complexity = 2, # "auto", 3, 2 or 1.
  use_ar1 = TRUE # default is FALSE
)
class(res)
sres <- tidy(res)
sres[["term"]] <- sub("gsp\\(.*\\)\\)", "gsp(...)", sres[["term"]]) # simplify output
sres
```

# Predicted Values

```{r}
ggplot() +
  aes(x = age, y = bmi) +
  geom_path(
    data = pheno_dt_female,
    mapping = aes(group = ID),
    alpha = 0.10,
    show.legend = FALSE
  ) +
  stat_smooth(
    data = pheno_dt_female,
    mapping = aes(colour = "On observed values"),
    se = FALSE,
    method = "gam",
    formula = y ~ s(x, bs = "cr")
  ) +
  stat_smooth(
    data = setnames(
      x = predict_bmi(res, start = min(pheno_dt_female[["age"]]), end = max(pheno_dt_female[["age"]])),
      old = c("egg_ageyears", "egg_bmi"),
      new = c("age", "bmi")
    ),
    mapping = aes(colour = "On predicted values"),
    se = FALSE,
    method = "gam",
    formula = y ~ s(x, bs = "cr")
  ) +
  scale_x_sqrt(
    expand = c(0, 0),
    breaks = c(0, 0.5, 1, 1.5, 2, 5, 10, 15),
    limits = c(0, NA)
  ) +
  scale_colour_manual(values = c("#b22222", "#2222b2")) +
  labs(
    x = "AGE (years)",
    y = "BMI (kg/m\u00B2)",
    colour = "GAM: y ~ s(x, bs = \"cr\")"
  ) +
  theme(
    legend.position = c(0.01, 0.99),
    legend.justification = c("left", "top"),
    legend.background = element_rect(fill = "white", size = 0.10),
    panel.grid = element_blank(),
    axis.line = element_line(size = 0.10)
  )
```

# Residuals

```{r}
plot_residuals(
  x = "age",
  y = "log(bmi)",
  fit = res
) +
  plot_annotation(
    title = "Cubic Splines (Random Linear Splines) - BMI - Female",
    tag_levels = "A"
  )
```

# Predicted Average Slopes

```{r}
res_pred_slopes <- egg_slopes(
  fit = res,
  period = c(0, 0.5, 1.5, 3.5, 6.5, 10, 12, 17)
)
head(res_pred_slopes)
```

```{r}
plot_egg_slopes(
  fit = res,
  period = c(0, 0.5, 1.5, 3.5, 6.5, 10, 12, 17)
)
```

# Area Under The Curve

```{r}
res_auc <- egg_aucs(
  fit = res,
  period = c(0, 0.5, 1.5, 3.5, 6.5, 10, 12, 17)
)
head(res_auc)
```

```{r}
plot_egg_aucs(
  fit = res,
  period = c(0, 0.5, 1.5, 3.5, 6.5, 10, 12, 17)
)
```
