---
title: "Run 'eggla' Workflow"
author: "Mickaël Canouil, *Ph.D.* (<***REMOVED***>)"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Run Non-Interactively

1. Copy and edit the following code to a new file (_e.g._, `run_eggla.sh`) on the server that will run the analysis with the appropriate parameters.

  - `renv` (recommended):

    ```{bash, eval = FALSE}
    #!/bin/bash

    home_analysis="/tmp/egg_analysis" # to be changed to the folder in which "egg_analysis" is to be performed

    mkdir $home_analysis 

    cd $home_analysis || exit

    Rscript \
      -e 'temp_library <- file.path(getwd(), "R")' \
      -e 'dir.create(temp_library, recursive = TRUE)' \
      -e '.libPaths(temp_library)' \
      -e 'install.packages("renv", lib = temp_library, repos = "http://cloud.r-project.org")' \
      -e 'library("renv")' \
      -e 'renv::init(bare = TRUE, settings = list(use.cache = FALSE))' \
      -e 'utils::download.file(
        url = "https://raw.githubusercontent.com/mcanouil/eggla/v0.5.2/inst/setup/renv.lock",
        destfile = "renv.lock"
      )' \
      -e 'renv::restore(lockfile = "renv.lock")' \
      -e 'renv::install("mcanouil/eggla@v0.5.2")' \
      -e 'unlink(temp_library, recursive = TRUE)'

    Rscript \
      -e 'library(eggla)' \
      -e 'library(data.table)' \
      -e 'res <- try(run_eggla(
        data = fread("/tmp/bmigrowth.csv"), # to be changed with the path of the file containing the data
        id_variable = "ID",
        age_days_variable = NULL, # computed based on "age_years_variable" if not provided. Only used for QC.
        age_years_variable = "age", 
        weight_kilograms_variable = "weight",
        height_centimetres_variable = "height",
        sex_variable = "sex",
        covariates = NULL,
        male_coded_zero = FALSE,
        parallel = FALSE, # to parallelise Daymont QC
        parallel_n_chunks = 1, # to parallelise Daymont QC
        working_directory = getwd() # or in that case "/tmp/egg_analysis"
      ))' \
      -e 'if (inherits(res, "try-error")) unlink(wd, recursive = TRUE)'
    ```

  - `pak` (faster):

    ```{bash, eval = FALSE}
    #!/bin/bash

    home_analysis="/tmp/egg_analysis" # to be changed to the folder in which "egg_analysis" is to be performed

    mkdir $home_analysis 

    cd $home_analysis || exit

    Rscript \
      -e 'temp_library <- file.path(getwd(), "R")' \
      -e 'dir.create(temp_library, recursive = TRUE)' \
      -e '.libPaths(temp_library)' \
      -e 'utils::install.packages(
        pkgs = "pak",
        lib = temp_library,
        repos = sprintf(
          "https://r-lib.github.io/p/pak/stable/%s/%s/%s",
          .Platform[["pkgType"]], R.Version()[["os"]], R.Version()[["arch"]]
        )
      )' \
      -e 'library(pak)' \
      -e 'utils::download.file(
        url = "https://raw.githubusercontent.com/mcanouil/eggla/v0.5.2/inst/setup/pkg.lock",
        destfile = "pkg.lock"
      )' \
      -e 'lockfile_install(lockfile = "pkg.lock", lib = temp_library)' \
      -e 'pkg_install("mcanouil/eggla@v0.5.2", lib = temp_library, upgrade = FALSE, dependencies = FALSE)'
    
    Rscript \
      -e 'library(eggla)' \
      -e 'library(data.table)' \
      -e 'res <- try(run_eggla(
        data = fread("/tmp/bmigrowth.csv"), # to be changed with the path of the file containing the data
        id_variable = "ID",
        age_days_variable = NULL, # computed based on "age_years_variable" if not provided. Only used for QC.
        age_years_variable = "age", 
        weight_kilograms_variable = "weight",
        height_centimetres_variable = "height",
        sex_variable = "sex",
        covariates = NULL,
        male_coded_zero = FALSE,
        parallel = FALSE, # to parallelise Daymont QC
        parallel_n_chunks = 1, # to parallelise Daymont QC
        working_directory = getwd() # or in that case "/tmp/egg_analysis"
      ))' \
      -e 'if (inherits(res, "try-error")) unlink(wd, recursive = TRUE)'
    ```

2.  Run the analysis in bash

    ```{bash, eval = FALSE}
    bash run_eggla.sh
    ```

3. Retrieve the two archives

    ```
    /tmp/egg_analysis/
    ├── 2021-11-23-female.zip
    └── 2021-11-23-male.zip
    ```

# Run Interactively

1. Create the working directory

    ```{bash, eval = FALSE}
    home_analysis="/tmp/egg_analysis" # to be changed to the folder in which "egg_analysis" is to be performed
    
    mkdir $home_analysis 

    cd $home_analysis || exit
    ```

2. Start R and setup working directory using `renv` (recommended) or `pak` (faster) to restore predefined version of packages

  - `renv` (recommended):

    ```{r, eval = FALSE}
    temp_library <- file.path(getwd(), "R")
    dir.create(temp_library, recursive = TRUE)
    .libPaths(temp_library)
    install.packages("renv", lib = temp_library, repos = "http://cloud.r-project.org")
    library("renv")
    renv::init(bare = TRUE, settings = list(use.cache = FALSE))
    utils::download.file(
      url = "https://raw.githubusercontent.com/mcanouil/eggla/v0.5.2/inst/setup/renv.lock",
      destfile = "renv.lock"
    )
    renv::restore(lockfile = "renv.lock")
    renv::install("mcanouil/eggla@v0.5.2")
    unlink(temp_library, recursive = TRUE)
    ```

  - `pak` (faster):

    ```{r, eval = FALSE}
    temp_library <- file.path(getwd(), "R")
    dir.create(temp_library, recursive = TRUE)
    .libPaths(temp_library)
    utils::install.packages(
      pkgs = "pak",
      lib = temp_library,
      repos = sprintf(
        "https://r-lib.github.io/p/pak/stable/%s/%s/%s",
        .Platform[["pkgType"]], R.Version()[["os"]], R.Version()[["arch"]]
      )
    )
    library(pak)
    utils::download.file(
      url = "https://raw.githubusercontent.com/mcanouil/eggla/v0.5.2/inst/setup/pkg.lock",
      destfile = "pkg.lock"
    )
    lockfile_install(lockfile = "pkg.lock", lib = temp_library)
    pkg_install("mcanouil/eggla@0.5.2", lib = temp_library, upgrade = FALSE, dependencies = FALSE)
    ```

3. Restart R

4. Run the analysis

    ```{r, eval = FALSE}
    library(eggla)
    library(data.table)
    res <- try(
      run_eggla(
        data = fread("/tmp/bmigrowth.csv"),
        id_variable = "ID",
        age_days_variable = NULL,
        age_years_variable = "age",
        weight_kilograms_variable = "weight",
        height_centimetres_variable = "height",
        sex_variable = "sex",
        covariates = NULL,
        male_coded_zero = FALSE,
        parallel = FALSE,
        parallel_n_chunks = 1,
        working_directory = getwd()
      )
    )
    if (inherits(res, "try-error")) unlink(wd, recursive = TRUE)
    ```

5. Retrieve the two archives

    ```
    /tmp/egg_analysis/
    ├── 2021-11-23-female.zip
    └── 2021-11-23-male.zip
    ```
